\documentclass{report}
\usepackage{a4}
\usepackage{fignum}

\newcommand{\amplot}{{\bf AMPlot2}}
\newcommand{\RA}[1]{\mbox{{\em Right-Amiga}--{\bf #1}}}
\newcommand{\degree}{\mbox{$\mbox{}^\circ$}}

\hyphenation{promp-ted}
\hyphenation{pro-ject}
\hyphenation{Post-Script}

\makeindex
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% COPYRIGHT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\Large\sf SciTech Software \bf Licence\index{Licence} Agreement
\end{center}
\vspace*{1ex}
This software is protected by United Kingdom Copyright\index{Copyright}
Law and International 
Treaty provisions. You must therefore treat this software like a book with the 
exception that you are authorised to make {\bf two} backup\index{Backing Up}
copies for the sole purpose of protecting your investment in this software.

The software is licensed for use on one Amiga computer at any one time by any
number of people. Thus, you 
may freely move the software from one computer to another as long as there is no 
possibility of the software being used at one location while it is being used at 
another location. Just as a book may only be read by one person at a time (unless 
it is illegally photocopied\ldots), this software may only be used by one person 
on a single computer at one time. If this software is in use on more than one 
computer at the same time, you are in breach of {\sf SciTech Software's} 
copyright and of this licence agreement. Consequently, you are breaking the law 
and forfeit the right to use this software.

\vspace*{1em}
\begin{center}
\Large\bf Warranty\index{Warranty}
\end{center}
\vspace*{1ex}
The physical diskette and documentation are warranted\index{warranty} to be free 
from defects in 
materials and workmanship for thirty (30) days from the date of purchase. If a 
defect in the diskette or the documentation is notified within 30 days, {\sf 
SciTech Software} will replace the defective diskette and/or documentation. Such 
warranty is limited to the replacement and shall not, in any event,
encompass any other damages, 
including, but not limited to, loss of profit, special, incidental, consequential, 
or other claims.

The program and documentation is sold ``as is'' and without warranty as to 
performance, merchantability, or fitness for a particular purpose. In particular,
while every effort has been made to make the PostScript output of this program
match the requirements of the PostScript and EPSF standards, no guarantee is made
that the software will work with all PostScript software. If you identify any
incompatibilities, you are encouraged to report them on the bug or enhancement
request forms supplied.

This warranty does not affect your statutory rights.

\vspace*{1em}
\begin{center}
\Large\bf Registration
\end{center}
\vspace*{1ex}
Please complete and return the registration\index{Registration}
form at the back of this manual.
Doing so will entitle you to updates to the software. Also supplied at the back 
of the manual and on the disk are Bug Report and Enhancement Request forms.

\tableofcontents

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% INTRODUCTION %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Introduction}
\amplot\ is a flexible scientific and technical graph plotting program
orignally written for the 
Commodore Amiga\footnote{Amiga, Workbench and Kickstart are trademarks of 
Commodore-Amiga, Inc.}
which produces PostScript\footnote{PostScript is a trademark of Adobe Systems, Inc.},
HPGL\footnote{HPGL is a trademark of Hewlett Packard, Inc.},
and (on the Amiga) IFF-DR2D output.

The Unix version of \amplot\ does not have a graphical interface ---
control of the program is \emph{via} control commands embedded in the datafile.

Unlike most graph-plotting programs on the Amiga, the emphasis in \amplot\ is on 
the quality of output---the screen representations of graphs are only 
approximations. \amplot\ uses a hires, interlaced screen\index{screen display}
on which a maximum of three
pen\index{pens} colours may be displayed (in addition to the background colour).

\amplot\ is designed to produce reproduction-quality graphs for use in scientific 
and technical documents. Four major graphing formats are supported: scatter plots, 
line plots, bar charts and pie charts. Within these major categories, numerous 
options are available.

On the Amiga, \amplot\ also has an extensive ARexx\footnote{ARexx is
Copyright Wishful Thinking Development Corp.} interface which allows
control over the graph, plotting of additional points and definition
of functions for fitting using the fitting routines included in
\amplot.

\section{Running \amplot}
Under Unix, \amplot\ is run using the command:
%\begin{verbatim}
%      amplot <datafile>
%\end{verbatim}
where \verb1<datafile>1 is the file of data to be plotted. Output is
to standard output, so you will wish to redirect this to a file or
pipe it to a PostScript printer or viewing program.

The file of data which is used as input to the program can be as simple as two 
columns of numbers or may contain a multitude of commands to describe the exact 
format of your graph (Chapter~\ref{ch:datafile}). You can create such a file with 
any text editor\index{editing}.



%%%%%%%%%%%%%%%%%%%%%%%%%% MENUS & REQUESTERS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\chapter{Menus and Requesters}
\label{ch:menus}
This chapter will describe each of the menus and requesters with its usage. The 
sections will be titled by menu name, with subsections matching menu items.

Under Unix, there are no Menus and Requesters (dialogue boxes); the
program is controlled \emph{via} commands in the data file (See
Chapter~\ref{ch:datafile}). This chapter has been left in place since
it describes the functions of those commands.

\section{The Project Menu}


\subsection{Paper}
{\bf Project/Paper}\index{Paper} displays a requester in which you can set 
the size of the output graph and, for PostScript and HPGL plots, its offset 
from the bottom left hand corner of the paper. PostScript assumes paper is in 
portrait orientation while HPGL assumes landscape orientation. These 
dimensions are specified in inches. The default paper size and position places 
the graph at the top of a sheet of A4 paper with a suitable offset at the left to 
allow the paper to be bound when using PostScript output. 
The dimensions actually refer to the bounds of the 
graph itself. The axis labels and titles will fall outside this area so the total 
size of the graph will be somewhat larger than the specified size.
In addition there is a checkbox for making the PostScript output conform to the
EPSF\index{EPSF} standard. When this gadget is selected, the Offset gadgets become 
redundant since the program which imports the EPSF file will handle offsets and image 
size. If the size of text on the graph is important, however, the sizes specified 
in this requester should match the final required graph size as closely as possible.


\subsection{Set Pens}
\amplot\ \index{Set Pens}has six `pens'. Each pen has an associated colour (for 
screen and HPGL output), thickness of line (for PostScript, EPSF and DR2D output), 
datapoint mark and join feature.
{\bf Project/Set Pens} displays a requester in which you can set the colour and 
thickness for each pen. Three colours (as well as the background colour) are 
available for display on the screen. Values greater than 3 may be specified if your 
HPGL plotter has more than 3 pens. The screen display will be restricted to one of 
the 3 screen pens. The line thickness is specified in points 
(1pt~$=\frac{1}{72}$inch). Your data may be assigned to any of the six pens. By 
default, the first set of data is displayed in Pen~1, the second in Pen~2, etc.
This may be changed using the {\bf Pen} item of the {\bf Lines} menu (See 
Subsection~\ref{ss:pen}).
However, the axes are always displayed in Pen~6 and the grid across the graph is 
always in Pen~5.



\section{The Style Menu}


\subsection{XY}
{\bf Style/XY} \index{XY}\index{Line Plot} is used to select the default 
line graph style. That is, each datapoint is joined by a line. 
The {\bf Style/Line Setup} menu item may be used to add marks to each 
datapoint. Each line may be displayed in any one of the six available pens (See 
Subsection~\ref{ss:pen}).


\subsection{Scatter}
The {\bf Style/Scatter} \index{Scatter Plot} graph style is similar to the {\bf XY} style, 
but, by default, the datapoints are not joined. Rather, they are marked, by default, 
with crosses. 
These defaults may be changed using the {\bf Style/Line Setup} menu item 
(Subsection~\ref{ss:dp}).


\subsection{Bar}
The {\bf Style/Bar} \index{Bar Chart} menu item displays your data as 
a bar chart. By default, each 
successive set of data is overlaid on the last set of data, but with the bars 
shrunk in width by 20\%. This shrink factor, together with other style options, 
such as outline, grouped and stacked charts may be set using the {\bf Style/Bar 
Style} menu option (Subsection~\ref{ss:barstyle}). You may fill the bars of 
your chart 
using the {\bf Fills} sub-items of the {\bf Lines} menu (Subsection~\ref{ss:fill}).
Note that the first bar will have a width dependent on the minimum X-boundary of 
the graph. Depending on your data, you may find it necessary to change the minimum 
x-boundary using the 
{\bf Set Axes} item of the {\bf Axes} menu (Subsection~\ref{ss:bounds}).
Alternatively, you may supply a datapoint in the graph with zero Y-value at a 
suitable X-value to achieve the correct bar width.

\subsection{Pie}
The {\bf Style/Pie} \index{Pie Chart} menu item displays your data as 
a Pie chart. This menu item will be disabled unless your datafile contains the 
keywords {\tt PIE} or {\tt COLUMNS 1}. Only one set of data may 
be displayed and a number of features of the other graph styles are not relevant. 
The `slices' of the pie may be filled using the {\bf Fills} item of the
{\bf Lines} menu (Subsection~\ref{ss:fill}) and may be ejected with the {\bf 
Style/Eject Slice} menu item. 


\subsection{Errors}
{\bf Style/Errors} (abbreviation: \RA{E}) allows \index{Errors} you to switch error 
bars on and off on your graph, providing 
error bar data has been supplied in your datafile.
The dataset may contain three columns of data, such that the same value is
used for both positive and negative parts of the
error bars or four columns of data specifying separate positive and
negative errors. The first value read (i.e.\ that in column 3 of the dataset)
is the negative error, the second value (i.e.\ that in column 4 of the dataset)
is the positive error. If any error value (positive or negative) is zero, no bar
will be drawn for this half of the error.

\subsection{Eject Slice}
{\bf Style/Eject Slice}\index{eject} allows slices of a Piechart to be 
ejected or replaced. 
Having selected this menu item, you must click on the pie slice to move. The 
percentage by which the slices are ejected is set in the {\bf Style/Pie Setup} 
requester.

\subsection{Line Setup}
\label{ss:dp}
{\bf Style/Line Setup} \index{Line Setup} (abbreviation: \RA{L}) displays a 
requester in which you may set the data point mark 
to be used with each pen in the XY and Scatter styles. Each pen also has a {\bf Link}
toggle gadget. If this is switched on for every pen, the Scatter and XY styles become 
equivalent. {\bf Link} causes datapoints to be joined by a line, even 
when the Scatter 
style has been chosen. This allows you to display one set of data simply as 
datapoint marks, with another set of data with the data points joined. This may 
be used if you have a set of experimental data which you wish to display as, for 
example, filled circles and a fitted curve which may be displayed with the join 
feature switched on, but with no datapoint marks. 

This requester also allows you to specify the width of the cross parts of error 
bars and the size of datapoint marks. Both of these values are specified in points.

\subsection{Bar Setup}
\label{ss:barstyle}
{\bf Styles/Bar Setup} \index{Bar Setup} displays a requester in which 
you can control the exact format of bar charts.

You may change the default (overlaid) barchart style to a grouped style 
\index{Bar Grouped} where successive sets of data are displayed adjacent to one 
another, or to a stacked style, where successive sets of data are stacked on top 
of eachother. There are certain restrictions on these 2 styles---you must have the 
same number of data points in each data set and the width of each bar must be the 
same. Normally, these styles are used for displaying numeric data on the Y-axis and 
largely non-numeric data (such as months, years, or names) along the X-axis. 
Thus, in the datafile, it is easiest to specify the X-axis data as 1,2,3\ldots\ 
switch off the X-axis labels with the {\bf No X-axis labels} gadget in the 
{\bf Axis \& Pie Labels} item of the {\bf Text} menu and supply your own X-axis 
labels using the same requester (Section~\ref{ss:bplab}).
Conversely the outline style \index{Bar Outline} is used when it is the area under 
the graph which is important and the bars may be of differing widths.

You may set the percentage by which successive sets of data are shrunk
\index{Bar Shrink} using the supplied gadget
and whether or not the first data set is shrunk using the {\bf Shrink First} 
checkbox. Bars are shrunk always shrunk in the grouped style; should you wish bars 
not to be shrunk, simply set the shrink factor to zero. 

The {\bf Fill all bars} checkbox controls whether fills are applied to the whole 
dataset or to individual bars.


\subsection{Pie Setup}
The {\bf Styles/Pie Setup}\index{pie setup} requester allows you to define where 
the first slice 
of the piechart is (0 = right-pointing x-axis, 90 = up-pointing y-axis, etc.).
Labels added to piecharts using the {\bf Text/Axis \& Pie Lables} requester will 
remain associated with the correct slice while labels added with the {\bf Text/Extra 
Labels} requester will not move as the starting angle of the piechart is changed.
This requester also allows you to specify the percentage by which slices are 
ejected from the pie.



\section{The Axes Menu}

\subsection{LogX}
{\bf Axes/LogX} causes \index{Log Axes} the X-axis to be displayed on 
a log scale (to base 10). Subticks will be 
placed between the labelled ticks. If your axis only
spans one log value (i.e.\ non-log values from 0.1 to 1, 1 to 10, 10 to 100, 
etc.), subticks will also be labelled if you switch on the {\bf Subticks} checkbox 
in the {\bf Axes/Set Axes} requester. \amplot\ will check that your data are valid 
for taking logs (i.e.\ all values are greater than zero). If the data are invalid, 
selecting this option will simply have no effect.


\subsection{LogY}
{\bf Axes/LogY} causes \index{Log Axes} the Y-axis to be displayed on 
a log scale (to base 10). Subticks will be 
placed between the labelled ticks. If your axis only
spans one log value (i.e.\ non-log values from 0.1 to 1, 1 to 10, 10 to 100, 
etc.), subticks will also be labelled if you switch on the {\bf Subticks} checkbox 
in the {\bf Axes/Set Axes} requester. \amplot\ will check that your data are valid 
for taking logs (i.e.\ all values are greater than zero). If the data are invalid, 
selecting this option will simply have no effect.



\subsection{Boxed}
{\bf Axes/Boxed} (abbreviation: \RA{B}) causes axis \index{Boxed Axes} 
lines to be placed to the top and right of the graph. The axes are drawn in Pen~6.


\subsection{Grid}
{\bf Axes/Grid} (abbreviation: \RA{G}) \index{Axis Grid} switches on a grid with 
grid lines along the tick mark positions. The grid is drawn in Pen~5.
When you select the grid, the boxed option will be turned on automatically.
If you have log axes, the grid will only be placed at the major tick positions. By 
switching on subticks in the {\bf Axes/Set Axes} requester, the grid will also be 
drawn at the subtick positions of log axes. The grid cannot be drawn at subtick 
positions of linear axes.

\subsection{Set Axes}
\label{ss:bounds}
{\bf Axes/Set Axes} (abbreviation: \RA{A}) displays a requester which allows you to 
define the minimum and maximum 
values displayed on each axis (bounds)\index{Bounds} and the spacing of ticks 
\index{Ticks} and subticks. 
If you change any of these, you may select the appropriate {\bf Auto} gadget to 
restore the program's calculated values.

The {\bf Clip} checkbox\index{clipping data} causes data outside the specified 
bounds not to be plotted 
while the {\bf Sub} checkbox switches on and off subticks on both axes. If you 
require subticks on one axis only, they should be switched on and the 
subtick spacing for the axis where they are not required should be set to 
0.0.
The {\bf Sub} checkbox also causes grid lines to be plotted at subtick positions of 
log axes and labels to be placed at subtick positions of log axes if they span only 
1 log unit.

If you enter a minimum value which is greater than the maximum value for that axis, 
the minimum and maximum values will be swapped automatically. If bounds are 
supplied for a log axis which span 0.0, a requester will be displayed asking you to 
enter positive bounds. Similarly, if your tick step size will generate more than 
1000 ticks, you will be asked to enter a larger value.

Ticks and subticks, like the axes, are drawn in Pen~6.

\subsection{Follow Zero}
{\bf Axes/Follow Zero} \index{Axis Follow Zero} causes the axes to follow 
the zero values on the graph, rather than being placed at the far left and bottom 
of the graph. If your graph does not span a zero value on both axes, a requester 
will appear to check this is really what you want to do.


\section{The Text Menu}

\subsection{Title}
{\bf Text/Title} \index{Title} displays a requester in which you may specify 
a title for the graph. 
The coordinates which you specify are the bottom, centre of the title (i.e.\ the 
title is centre, bottom justified). In all requesters which require graph coordinates
to be specified you may click in the main graph window rather than typing in numbers.
When you click, the coordinates will be transferred into the requester.
By default the title will be centred on the 
graph at 90\% of its height. Choosing the {\bf Centre} gadget will restore this 
default position. Once a title has been specified with its coordinates, the title 
will be locked to these coordinates. Thus, if you change the bounds of the graph, 
the title will move. You will then need to return to this requester to restore 
the position on the graph where you wish the title to appear. Clearing the title 
string gadget will switch the title off. You may also specify the name and size
of the font to be used. (See Chapter~\ref{ch:fonts}.)


\subsection{Axis Titles}
{\bf Text/Axis Titles} \index{Axis Titles} displays a requester in which you can 
enter titles for the axes. The requester also lets you specify the fonts to be used
(Chapter~\ref{ch:fonts}).
The labels will be centred along the axes. Although the Y-axis title is displayed 
down the screen, on paper, the label will be displayed sideways, reading up the paper.


\subsection{Axis \& Pie Labels}
\index{Axis Labels}
This requester allows you to set the fonts used for the axis labels;
the font specified for X-axis labels will also be used for labels placed next 
to slices of a piechart. 
The precision of the axis labels may also be set. This is the number of decimal 
places to be displayed. If a negative value is given, exponential form is used for 
the labels. If a label will require more than 24 characters to print, it will 
automatically be switched to exponential form. For log axes, only the sign of the 
precision is taken into account---if positive, the precision will be 0 for label 
values greater than 1.0 and will be adjusted automatically for label values less 
than 1.0; if negative, exponential form will be used with 1 decimal place.

The three check boxes allow labelling of the axes to be switched off and labels 
on the x-axis to be centred. This can be useful for barcharts.

Clicking the {\bf Bar/Pie Labels} gadget brings up a further Bar/Pie Labels requester.

\subsubsection{Bar/Pie Labels Requester}
\label{ss:bplab}
This requester allows you to specify text for each bar or pie slice. You 
can step through the labels with the {\bf Next} and {\bf Prev} gadgets. The 
{\bf Remove} gadget removes all the labels. For datasets containing more than 100 
items, this requester 
will not be displayed since it would not be possible to place all the 
labels along the x-axis or to see all the labels crowded around a piechart.

With piecharts, the exact position of the label depends on the quadrant of the graph 
into which the label is being placed. (The first quadrant starts from the 
right-pointing horizontal axis, the second quadrant starts at the upward-pointing 
vertical axis, etc.)

\begin{center}
\begin{tabular}{|c|l|}\hline
Quadrant & Justification   \\ \hline
1        & Left            \\
2        & Right           \\
3        & Right           \\
4        & Left            \\ \hline
\end{tabular}
\end{center}

If you require more precise placing of labels you may use the {\bf Text/Extra Labels}
requester described below to place arbitrary additional labels, but note that these 
will not move with the piechart if you change the start angle with {\bf Style/Pie 
Setup}.


\subsection{Key}
{\bf Text/Key}\index{key} displays a requester which allows you to specify 
some text for 
each dataset to create a key. As with the {\bf Bar/Pie Labels} requester you can 
step through the labels with the {\bf Next} and {\bf Prev} gadgets. The font for 
the key is specified here (see Chapter~\ref{ch:fonts}) and the position of the 
first item must also be specified.
This may be done by clicking on the graph. The {\bf Remove} gadget will switch off 
the key. The key will appear as a set of lines and datapoints, or (for barcharts 
and piecharts) a set of appropriately shaded swatches next to the text.


\subsection{Extra Labels}
\label{ss:linlab}
{\bf Text/Extra Labels} \index{Extra Labels} displays a requester in which you may 
add arbitrary labels to your graph.
You must specify the label text, coordinates (bottom, left justified), 
font and font size (see Chapter~\ref{ch:fonts}).
The position of the label may be specified simply by clicking in the main graph 
window.
To change which label is being processed, click on the {\bf Next} or {\bf Prev}
gadgets. You may add as many labels as you like (subject to available memory) and 
may remove labels by clicking on the {\bf Kill} gadget. The {\bf Kill All} gadget 
removes all labels.



\section{The Lines Menu}


\subsection{Pen}
\label{ss:pen}
{\bf Lines/Pen} \index{Pens} allows you to alter the default mapping of pen to 
dataset. After typing a
pen number into the text gadget and selecting the {\bf OK} gadget you should click
on a datapoint of a line (you should click on a datapoint 
and not in the middle of a line). This data set will then be displayed in your chosen 
pen.


\subsection{Style}
The subitems of {\bf Lines/Style} \index{Line Style} allows you to set the line 
style for a dataset. This is independent 
of the pen, so you may have the same pen in solid and dotted line styles. Note 
that the PixelScript PostScript interpreter (V1.1) does not support dashed line 
styles. Note that the exact pattern of lines in an HPGL plot may depend on your 
plotter.


\subsection{Fills}
\label{ss:fill}
Bar charts \index{Bar Fills} and pie charts \index{Pie Fills} may be filled 
using one of 6 shades of grey or hatch styles available 
through this option. This gives a total of 7 shades including white. It is simply 
necessary to select a fill and click within a bar or slice of a pie chart. The 
increasing darknesses  of fills equate to the following hatch patterns:
\begin{center}
\begin{tabular}{|l|l|} \hline
Shade &  Hatch                      \\ \hline
1     &  Upward slant               \\
2     &  Downward slant             \\
3     &  Horizontal                 \\
4     &  Vertical                   \\
5     &  Horizontal \& Vertical     \\
6     &  Upward \& Downward slant   \\ \hline
\end{tabular}
\end{center}
The {\bf Lines/Fill Control} requester allows you to switch between shaded and 
hatched fills.


\subsection{Fill Control}
This\index{fill control,hatching,shading} allows you control over fills within the 
limitations of the display 
device. PostScript, EPSF and IFF-DR2D output allow both hatched and 
shaded fills, while HPGL allows only hatched fills. The screen only shows 
shaded fills. The thickness (for PostScript and IFF-DR2D) and spacing of 
the hatching may also be specified. You cannot mix hatched and shaded fills in a 
single plot. Note that hatching will have different effects with HPGL compared with 
IFF-DR2D and PostScript. Because of the nature of pen plotters, hatches will be 
overlayed in HPGL, while IFF-DR2D and PostScript will fill each area with white 
before hatching.

\subsection{No Fills}
{\bf Fills/No Fills} removes all fills.

\section{The Fitting Menu}


\subsection{Regression}
{\bf Fitting/Regression} \index{Regression} displays a requester in 
which you may specify a degree for linear regression polynomial fitting. 
This fits an equation of the form:
$$y=a+bx+cx^2+dx^3+ex^4\ldots$$
to your data using `singular value decomposition' (SVD) by minimising a $\chi^2$ 
merit function.
The polynomial degree which you specify sets the number of terms in 
the fitted equation. A degree of 2 will fit a straight line. You may specify 
degrees between 2 and 20, but the degree must not be greater than the number of 
datapoints in a dataset. If this is so, a requester will be posted warning you that 
the polynomial degree is too high.
The fitting routine assumes that the all the datapoints have the same
standard deviation. The coefficients derived for the fit will be displayed in the
message window in the order shown in the equation above.

The tolerence specifies the accuracy to which the fitting is performed. If 
the coefficients are of the same order as the tolerence, a much smaller 
value should be specified for the tolerence. The {\bf Write Data} checkbox 
causes detailed information on the fitting to be written to a file called 
{\tt
AMPlotFit.data} in the directory from which \amplot\ was run. This file contains the 
coefficients, variance and covariance matrix for each dataset. 

Selecting the {\bf Off} gadget will switch the fitted curve off.



\subsection{Robust}
{\bf Fitting/Robust} switches \index{Robust Fitting}
robust straight line fitting on or off. This form of 
fitting weights against values lying outside the normal distribution of values 
observed by the criterion of least absolute deviations. 
For the equation,
$$y = a + bx$$
this function minimises the merit function:
$$\sum_{i=1}^{N}\mid y_i-a-bx_i\mid$$
rather than the $\chi^2$ function minimised in normal linear regression.


\subsection{Fourier}
{\bf Fitting/Fourier} \index{Fourier Smoothing} 
displays a requester in which you may set a 
factor for Fourier transform smoothing to remove noise from your data.
A fast fourier transform (FFT) is 
performed, high frequency oscillations in the data are removed using the smoothing 
factor, and an inverse-FFT is performed on the filtered data.
Smoothing factors greater 
than half the number of data points render the graph virtually feature-less.
The values need not be integers.



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% DATAFILE %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{The \amplot\ Datafile}
\label{ch:datafile}
\index{Datafile}\index{Commands}
Almost every feature available from the menus may also be set from within 
the datafile using simple command keywords. All the available commands are shown 
in Table~\ref{tab:commands}. 
The commands may be in upper or lower case and may be abbreviated to the 
fewest letters necessary for the commands to be unambiguous. In all cases 
where text is required, if more than one word is to be included,
the text must be enclosed in double inverted commas. For example, when 
setting the title of a graph:
\begin{verbatim}
      Title "A Graph Title" 0.0 0.0
\end{verbatim}

Table~\ref{tab:commands} 
shows command applicability in an easy-to-read format.

\begin{center}
\scriptsize
\begin{tabular}{lccc} \hline
                                 &           &        & Commands which     \\
                                 & \multicolumn{2}{c}{Commands used in} %
                                                      & do not cause the   \\ \cline{2-3}
Command                          & Datafile  & ARexx  & graph to replot.   \\ \hline
{\bf Data}                       &           &        &                    \\
COLUMNS  n                       &     x     &        &                    \\
NDATA  n                         &     x     &        &                    \\
NEWSET                           &     x     &        &                    \\ \hline
{\bf Output}                     &           &        &                    \\
EPSF                             &     x     &     x  &       x            \\
OFFSET  x   y                    &     x     &     x  &       x            \\
SIZE  x   y                      &     x     &     x  &       x            \\ \hline
{\bf Pens}                       &           &        &                    \\
PALETTE  scrpen   r   g  b       &     x     &     x  &       x            \\
PEN  pen                         &     x     &     x  &       x            \\
SETPEN  pen   colour   thickness &     x     &     x  &                    \\
STYLE  style                     &     x     &     x  &       x            \\ \hline
{\bf Styles}                     &           &        &                    \\
BARCHART                         &     x     &     x  &                    \\
DATAPOINT  pen   mark   join     &     x     &     x  &                    \\
EJBY  percent                    &     x     &     x  &                    \\
EJECT  slicenum                  &     x     &     x  &                    \\
ERRORS                           &     x     &     x  &                    \\
EWIDTH  width                    &     x     &     x  &                    \\
GROUPED                          &     x     &     x  &                    \\
HATCH                            &     x     &     x  &       x            \\
HSPACE  space                    &     x     &     x  &       x            \\
HTHICK  thick                    &     x     &     x  &       x            \\
MARKSIZE  size                   &     x     &     x  &                    \\
NOERRORS                         &     x     &     x  &                    \\
OUTLINE                          &     x     &     x  &                    \\
OVERLAYED                        &     x     &     x  &                    \\
PIE                              &     x     &     x  &                    \\
PSTART  degrees                  &     x     &     x  &                    \\
SCATTER                          &     x     &     x  &                    \\
SHADE                            &     x     &     x  &       x            \\
SHRFIRST                         &     x     &     x  &                    \\
SHRINK  val                      &     x     &     x  &                    \\
STACKED                          &     x     &     x  &                    \\
XY                               &     x     &     x  &                    \\ \hline
\end{tabular}
\normalsize
\end{center}
Table~\ref{tab:commands} Continued over\ldots

\begin{center}
\scriptsize
\begin{tabular}{lccc} \hline
                                 &           &        & Commands which     \\
                                 & \multicolumn{2}{c}{Commands used in} %
                                                      & do not cause the   \\ \cline{2-3}
Command                          & Datafile  & ARexx  & graph to replot.   \\ \hline
{\bf Fills}                      &           &        &                    \\
FILL  num   x   y                &     x     &     x  &                    \\
NOFILLALL                        &     x     &     x  &                    \\
PFILL  num   angle               &     x     &     x  &                    \\ \hline
{\bf Axes}                       &           &        &                    \\
BOUNDS xmin xmax ymin ymax       &     x     &     x  &                    \\
BOXED                            &     x     &     x  &                    \\
CBOUNDS xmin xmax ymin ymax      &           &     x  &                    \\
CLIP                             &     x     &     x  &                    \\
FZERO                            &     x     &     x  &                    \\
GRID                             &     x     &     x  &                    \\
LOGX                             &     x     &     x  &                    \\
LOGY                             &     x     &     x  &                    \\
NOCLIP                           &     x     &     x  &                    \\
SUBOFF                           &     x     &     x  &                    \\
SUBON                            &     x     &     x  &                    \\
SUBTICKS  x   y                  &     x     &     x  &                    \\
TICKS  x   y                     &     x     &     x  &                    \\ \hline
%  \end{tabular}
%  \normalsize
%  \end{center}
%  Table~\ref{tab:commands} Continued over\ldots
%  
%  \begin{center}
%  \scriptsize
%  \begin{tabular}{lccc} \hline
%                                   &           &        & Commands which     \\
%                                   & \multicolumn{2}{c}{Commands used in} %
%                                                        & do not cause the   \\ \cline{2-3}
%  Command                          & Datafile  & ARexx  & graph to replot.   \\ \hline
{\bf Labels and Titles}          &           &        &                    \\
BPLAB  n   text                  &     x     &     x  &                    \\
CENTXLAB                         &     x     &     x  &                    \\
KEYFONT  font   size             &     x     &     x  &                    \\
KEYPOS  x   y                    &     x     &     x  &                    \\
KEYTEXT  n   text                &     x     &     x  &                    \\
LABEL n str x y font size        &     x     &     x  &                    \\
LFONT  font   size               &     x     &     x  &                    \\
NOCENTXLAB                       &           &     x  &                    \\
NOXLAB                           &     x     &     x  &                    \\
NOYLAB                           &     x     &     x  &                    \\
PRECISION  xprec   yprec         &     x     &     x  &                    \\
TFONT  font   size               &     x     &     x  &                    \\
TITLE  str   x   y               &     x     &     x  &                    \\
XLABEL  str                      &     x     &     x  &                    \\
YLABEL  str                      &     x     &     x  &                    \\
XLFONT  font   size              &     x     &     x  &                    \\
YLFONT  font   size              &     x     &     x  &                    \\
XTFONT  font   size              &     x     &     x  &                    \\
YTFONT  font   size              &     x     &     x  &                    \\ \hline
\end{tabular}
\normalsize
\end{center}
Table~\ref{tab:commands} Continued over\ldots

\begin{table}[h]
\begin{center}
\scriptsize
\begin{tabular}{lccc} \hline
                                 &           &        & Commands which     \\
                                 & \multicolumn{2}{c}{Commands used in} %
                                                      & do not cause the   \\ \cline{2-3}
Command                          & Datafile  & ARexx  & graph to replot.   \\ \hline
{\bf Fitting}                    &           &        &                    \\
FOURIER  n                       &     x     &     x  &                    \\
REGRESS  n                       &     x     &     x  &                    \\
REGWRITE                         &     x     &     x  &       x            \\
REXXFIT  n   path/macro          &     x     &     x  &                    \\
REXXWRITE                        &     x     &     x  &       x            \\
ROBUST                           &     x     &     x  &                    \\
RTOL  n                          &     x     &     x  &                    \\
TOL  n                           &     x     &     x  &                    \\ \hline
{\bf Miscellaneous Commands}     &           &        &                    \\
CBOUNDS xmin xmax ymin ymax      &           &     x  &                    \\
DRAW  x   y  [x   y]             &           &     x  &                    \\
MACRO  path/macro                &     x     &        &                    \\
MARK  x   y   num                &           &     x  &                    \\ \hline
\end{tabular}           
\end{center}
\caption[Datafile and ARexx commands]{\label{tab:commands}Datafile and ARexx 
commands. Crosses 
in the appropriate columns indicate whether commands are available in a 
datafile or through ARexx. The final column indicates ARexx commands 
which do not cause the graph to replot.}
\end{table}
\clearpage

\section{Data control commands}
\index{Commands: Data control}
\begin{description}
\item[!] This is used to introduce a comment either at the beginning of a 
line or after a command and its parameters.
\item[NEWSET] This is used to introduce a new set of data, for multi-line 
graphs. Optionally, {\tt NEWSET} may be used before the first set of data.
\item[COLUMNS] This is followed by a single parameter representing the 
number of columns of data to be read. In most cases this will be 2 (which is the 
default) i.e.\ columns representing X and Y. If set to 1, the Y values will all be 
set to zero. This is used only for Pie charts. If set to 3, the third column of data 
will be used for the size of $\pm$ error bars. Thus the total size of each error 
bar will be twice this value. If set to 4, separate error values will be read for
the $+$ and $-$ error bars. The first value read (i.e.\ that in column 3) is the
negative error and the second value read (i.e.\ that in column 4) is the positive
error.
\item[NDATA] The {\tt NDATA} command is followed by a parameter used to specify 
the number of datapoints which may 
follow. This defaults to 100. If a dataset has more than 100 points, the {\tt NDATA}
command should be used to specify a greater maximum number. Alternatively, if you 
only have a few datapoints, you can set a value of {\tt NDATA} less than 100 
to conserve memory.
\end{description}

\section{Output}
\index{Commands: Output}
\begin{description}
\item[SIZE] This is followed by the X and Y dimension of your output graph
in inches. Note that this is the size of the axes; any axis  labels, titles, etc.\ 
will fall outside these dimensions. EPSF and IFF-DR2D plots may be scaled by the 
program which imports them.
\item[OFFSET] This is followed by the X and Y offset of the bottom left hand 
corner of the graph in inches. This command is ignored if you are producing EPSF or 
IFF-DR2D output.
\item[EPSF] Sets an output PostScript file to conform to the EPSF standard.
\end{description}

\section{Pens}
\index{Commands: Pens}
\begin{description}
\item[PEN] This is followed by the pen number (1--6) to be used for the current
dataset. If there is more than one dataset, this command should come after the
{\tt NEWSET} command (though it need not come immediately after).
\item[STYLE] This is followed by a line style (1--6) to be used for the current
dataset. If there is more than one dataset, this command should come after the
{\tt NEWSET} command (though it need not come immediately after).
\item[SETPEN] This is followed by the pen number (1--6), the colour (between  1 and 
the number of pens you have for HPGL plotting; 1--3 if you are not interested in HPGL 
output) and the thickness for the PostScript output (in points).
\item[PALETTE] This is followed by the screen pen number (0--3) and the R, G and B 
values for the screen pen (each between 0 and 15).
\end{description}

\section{Styles}
\index{Commands: Graph style}
\label{sec:styles}
\begin{description}
\item[XY] Specifies the default line graph style.
\item[SCATTER] Specifies a scatter plot.
\item[BARCHART] Specifies a barchart.
\item[PIE] Specifies a pie chart.
\item[ERRORS] Switch on error bars if a 3 column dataset has been used.
\item[NOERRORS] Switches off error bars (default).
\item[DATAPOINT] This is followed by the pen number (1--6), the datapoint mark
to be used by this pen (0--12) and a flag to indicate if this set of points should 
be joined (1) or not (0). The mark numbers are specified as follows:\\
0: no mark\\
1: cross\\
2: square\\
3: filled square\\
4: circle\\
5: filled circle\\
6: diamond\\
7: filled diamond\\
8: triangle\\
9: filled triangle\\
10: inverted triangle\\
11: filled inverted triangle\\
12: plus sign
\item[EWIDTH] This is followed by the width of the cross part of an error bar (in 
points). The pen and hence the width of the lines which form the error bars is the 
same as that used for the current dataset.
\item[MARKSIZE] This is followed by the size of datapoint marks (in points).
\item[SHRINK] This is followed by the bar shrinkage percentage used for overlayed 
and grouped bars.
\item[SHRFIRST] Specifies that all sets of bars should be shrunk in the overlayed 
style.
\item[OVERLAYED] Specifies the overlayed bar style (default).
\item[OUTLINE] Specifies the outline bar style.
\item[GROUPED] Specifies the grouped bar style.
\item[STACKED] Specifies the stacked bar style.
\item[EJECT] This is followed by  a slice number for a slice to be ejected from a 
piechart. The slices are numbered from 1.
\item[EJBY] This is followed by the percentage of the piechart diameter by which 
slices should be ejected.
\item[PSTART] This is followed by the angle at which a piechart should start. This 
defaults to 0\degree\ which is the right-pointing horizontal axis. 90\degree\ 
specifies up-pointing vertical axis, etc.
\item[HATCH] This specifies that fills should be hatched. This is the default for 
HPGL plots where shading cannot be used. Note that the screen display shows only 
shading.
\item[SHADE] This specifies that fills should be shaded. This is the default except 
in the case of HPGL plots where shading is not possible and this command will be 
ignored.
\item[HTHICK] This is followed by the hatch fill line thickness for PostScript, EPSF 
and IFF-DR2D output. The value is specified in points.
\item[HSPACE] This is followed by the hatch fill line spacing. The value is 
specified in points.
\end{description}

\section{Axes}
\index{Commands: Axes}
\begin{description}
\item[NOXLAB] Switches off the labels on the X-axis.
\item[BOXED]  Boxes the axes. The axes are drawn in Pen~6.
\item[LOGX] Take logs (to base 10) of the X-axis data.
\item[LOGY] Take logs (to base 10) of the Y-axis data.
\item[BOUNDS] This is followed by the boundary values: XMin XMax YMin and YMax.
\item[CBOUNDS\dag] This is followed by XMin, XMax, YMin and YMax values. This works 
in much the same way as the {\tt BOUNDS} command, but the `best' clean bounds are 
calculated on these values. (This command may not work\ldots)
\item[TICKS]  This is followed by the tick step size along X and Y.
\item[SUBTICKS]  This is followed by the subtick step size along X and Y.
\item[SUBON] Switches on axis subticks.
\item[SUBOFF] Switches off axis subticks.
\item[GRID] Switches on the grid across the graph. The grid is drawn in Pen~5.
\item[FZERO] Causes the axes to follow zero on the graph, rather than being 
placed along the bottom and left of the graph.
\item[CLIP] Causes graph lines to be clipped at the bounds. (Default.)
\item[NOCLIP] Causes graph lines not to be clipped at the bounds.
\end{description}

\section{Labels and Titles}
\index{Commands:Text}
\begin{description}
\item[XLABEL] This is followed by a label for the X-axis. If this is more than 
one word, it should be enclosed in double inverted commas.
\item[YLABEL] This is followed by a label for the Y-axis. If this is more than 
one word, it should be enclosed in double inverted commas.
\item[LFONT] This is followed by a font and point size which will be used for X and Y 
axis labels and titles. This command is retained for compatibility with {\bf AMPlot} 
V1.0. \amplot\ now has separate commands to set the font for each of these items.
\item[XTFONT] This is followed by the font and point size for the X axis title.
\item[YTFONT] This is followed by the font and point size for the Y axis title.
\item[XLFONT] This is followed by the font and point size for the X axis labels.
\item[YLFONT] This is followed by the font and point size for the Y axis labels.
\item[TITLE] This is followed by a string for the graph title (in double inverted 
commas if more than one word) and the X and Y coordinates of the bottom centre of 
the string. 
\item[TFONT] This is followed by the font and point size for the title.
\item[LABEL] This is followed by a label number, a string (in double inverted 
commas if more than one word), the X and Y coordinates of the label, the font
name and font point size. 
\item[NOXLAB] Switches off labels on the X-axis.
\item[CENTXLAB] Centres x-axis labels. Used mainly for barcharts.
\item[NOCENTXLAB\dag] Switches off centering of X-axis labels. (Default.)
\item[NOYLAB] Switches off labels on the Y-axis.
\item[PRECISION] This is followed by the precision for the X and Y-axes. If the 
values are positive, the labels will be displayed in normal numeric format with the 
number specifying the number of decimal places. If negative, numbers will be 
displayed in exponential  form. With log axes, only the sign of the precision is 
important.
\item[BPLAB] This is followed by a number for the bar or pie slice to be labelled
and by the text for that label. If this is more than one word, it should be enclosed 
in double inverted commas. The bars or pie slices are numbered from 1. When 
specified in a datafile, this command will 
automatically switch on bar labels for display. Note that bar/pie labels are not 
available if any dataset contains more than 100 items.
\item[KEYPOS] This is followed by the X and Y position for the start of the key.
\item[KEYFONT] This is followed by the font and point size for the key.
\item[KEYTEXT] This is followed by a dataset number and the text for the key for 
this dataset. If the text contains more than one word, it should be enclosed 
in double inverted commas.
\end{description}

\section{Fills}
\index{Commands: Fills}
\begin{description}
\item[NOFILLALL] Do not fill all bars in a dataset. i.e.\ a fill must be 
specified for each bar in turn.
\item[FILL] This is followed by the fill shade (1--6, light to dark)
and X and Y coordinates 
about which to perform a fill. This command is not used for Pie charts, where
{\tt PFILL} should be used.
\item[PFILL] This is followed by the fill shade (1--6, light to dark)
and an angle (in degrees).
{\tt PFILL} is used only for Pie charts and the slice of pie at this angle will be 
filled. The angle is defined in the standard manner as the anti-clockwise angle
from the X-axis (i.e.\ the right-pointing  horizontal axis).
\end{description}

\section{Fitting}
\index{Commands: Fitting}
\begin{description}
\item[REGRESS] This is followed by an integer between 2 and 20 to set the degree 
of the polynomial to be used.
\item[TOL] This is followed by the tolerence to be used in the regression fitting.
\item[REGWRITE] Causes fitting data to be written to the file\\
{\tt AMPlotFit.data}.
\item[ROBUST] Switches on robust straight line fitting.
\item[FOURIER] This is followed by a smoothing factor for use in Fourier 
smoothing.
\item[REXXFIT] This is followed by an integer between 2 and 20 to set the degree 
of the ARexx polynomial to be used and the name of the function macro. This must 
include the full device and path name for the macro.
\item[RTOL] This is followed by the tolerence to be used in the ARexx regression 
fitting.
\item[REXXWRITE] Causes ARexx fitting data to be written to the file 
{\tt AMPlotRexxFit.data}.
\end{description}

\section{Miscellaneous Commands}
\index{Commands: Miscellaneous}
\begin{description}
\item[MARK\dag] This is followed by the X and Y-coordinates at which a datapoint mark 
should be placed and the mark style number as described in Section~\ref{sec:styles} 
for the {\tt DATAPOINT} command.
\item[DRAW\dag] This is followed by 2 or 4 parameters. If 2 parameters are given,
a line is 
drawn from the current position to these coordinates. If 4 parameters are given they 
are the start coordinates and end coordinates of a line. If the start coordinates 
are identical to the end coordinates, the command is treated as a `move' rather than 
a `draw' command and the current coordinates are simply updated without drawing 
anything onto the graph. The {\tt PEN} command may be given first to define the pen 
to be used.


\end{description}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% FONTS %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\chapter{Fonts and Text}
\label{ch:fonts}

\section{PostScript Fonts}
You\index{PostScript,Fonts} may specify any font you have available on 
your printer. Typical fonts and styles are shown in Table~\ref{tab:fonts}. 
PostScript is case sensitive when specifying font names. However, \amplot\ will
automatically capitalise the first letter of the font name, the first letter after a 
`--' and of the words Roman, 
Bold, Italic and Oblique. In addition, you may specify `Times' rather than
`Times--Roman'; the `--Roman' will be added by the program.
The\index{fonts} Amiga will display any font it doesn't know as Times.
You may use Amiga fonts other than Times and Helvetica, although, unless you have 
fonts of these names on your PostScript printer, you will not be able to print 
them.
\begin{table}
\begin{center}
\begin{tabular}{|l|l|} \hline
Helvetica                & Times--Roman        \\
Helvetica--Bold          & Times--Bold         \\
Helvetica--Oblique       & Times--Italic       \\
Helvetica--BoldOblique   & Times--BoldItalic   \\ \hline
\end{tabular}
\end{center}
\caption[Typical PostScript Fonts]{\label{tab:fonts}Typical PostScript fonts 
available on most PostScript printers and emulators. These are Adobe fonts and 
non-Adobe printers and emulators may have lookalike fonts, though typically they 
will also recognise these names.}
\end{table}

\section{Font Translation}
\amplot\ provides a font translation table\index{font translation} for specifying 
screen fonts and HPGL\index{HPGL}
fonts. Two examples of the translation table are provided: {\tt amplot.ftrans.V2.0} 
and {\tt amplot.ftrans.V1.3}. The install procedures will copy the appropriate 
version to {\tt amplot.ftrans} in your {\tt S:}\ directory. When \amplot\ starts it 
looks for {\tt amplot.ftrans} first in your current directory, then in the {\tt S:}\ 
directory. The file is read only on program startup; if you alter the file, you must 
restart \amplot.

The file is free format and contains three columns. Each line must contain the 
following three pieces of information.
\begin{itemize}
\item A PostScript or IFF-DR2D fontname.
\item The equivalent screen font.
\item The equivalent HPGL font number.
\end{itemize}
Using this table, you can select the best screen font to display for each PostScript 
or IFF-DR2D font name. Note that the screen fonts are specified without the 
{\tt .font} part of the name. Fonts in HPGL are simply specified by number. 
Thus, you must also give a number for HPGL output. You should try to select an HPGL 
font which looks something like the PostScript font.

The table also provides a further important function. You should specify a 
PostScript font name of `Alternate'. This is followed not by an equivalent screen 
font, but by your chosen alternate font and HPGL font number. The alternate font is 
normally used for Greek characters and mathematical symbols and you would normally 
select the PostScript {\bf Symbol} font. The following section explains how you 
should access the alternate font.

If a font is not found in the translation table, a simple translation scheme is 
applied which just drops any part of the font name after a {\bf .}, or {\bf --}.
If no Amiga font is found with this name, the Times font is used.


\section{Special Characters}
\amplot\ allows\index{special characters} you to introduce two types of special 
characters into any piece of 
text which you can control directly (i.e.\ titles, extra labels, etc.). Note that 
these special characters will {\em not\/} appear on the screen display; the normal 
Roman style character will appear.

The first 
type or character is a superscript\index{superscript} or subscript\index{subscript}.
To produce a superscript, use a 
\verb1^1; to produce a subscript, use a \verb1_1. If you need to raise or lower a 
group of characters, enclose them in curly brackets. For example, the Y-axis title of 
your graph might be:
$$ y = a + bx_1 + cx^{12} $$
To obtain this label, you should enter the following title into the {\bf Text/Axis 
Titles} requester:
\begin{verbatim}
                  y = a + bx_1 + cx^{12}
\end{verbatim}

The second type of character is one from an alternate font. Typically, you might 
wish to include Greek\index{Alternate font, Greek} characters. These are supplied 
in the PostScript {\bf Symbol} 
font. To obtain characters from the alternate font, you should precede  the normal 
character with a backslash. The exact mapping of normal Roman style font characters 
to the alternate font will depend on the font you select. If we assume that the 
letter `a' will generate the character $\alpha$ in the alternate font and you wish 
to produce the label:
$$ y = 3\alpha x + x^2 $$
you would enter the label as:
\begin{verbatim}
                  y = 3\ax + x^2
\end{verbatim}
Note that you must place a backslash before every character to be in the alternate 
font; you cannot group characters using curly brackets.

The font translation table described in the previous section is used to specify the 
alternate font.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% HISTORY %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\chapter{History of \amplot}
\amplot\ was inspired by a simple graph plotting program named CURVY written by 
Dr.\ Andrew J.\ McLachlan of the M.R.C.\ laboratory in Cambridge. CURVY was 
written in FORTRAN for the VAX computer using a plotting library called PLOT84.
The first version of \amplot\ was written in 1989 to accept data files in the 
same format used by CURVY, processing a subset of its commands, but producing 
PostScript output. The original version of {\bf AMPlot} was written in generic
portable C on the Amiga, so that the program could be compiled and run on any 
computer. During 1990, the menu interface, full command interface and additional 
features such as bar charts, pie charts, curve fitting routines and datapoint marks 
were added. \amplot\ was a complete re-write performed in 1992 adding many new 
features including HPGL and DR2D output, ARexx support and the 3D-look
interface.

In 1997, a Unix version was produced which lacks the graphical interface.

The 3D-look interface was written using the {\bf STSLib} gadget and menu library 
which is also available from {\sf SciTech Software}.

\amplot\ was developed by Dr.\ Andrew Martin on an Amiga 3000/25/100 computer.
The manual was produced using Amiga\TeX\ (Radical Eye Software) and
\LaTeX. Porting to Unix was performed under Linux.


\newpage
\input{amplot.index}
\end{document}
